---
title: "EID_and_EVT_summarydoc.Rmd"
author: "Gustaf Rydevik"
date: "3 August 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
script.path<-"./scripts"
autolib<-function(Package,mirror="http://stat.ethz.ch/CRAN",...){
  if(!suppressWarnings(require(as.character(substitute(Package)),
                               character.only=TRUE,quietly=TRUE,...))){
    Package<-as.character(substitute(Package))
    install.packages(Package,repos=mirror,...)
    library(Package,character.only=TRUE)
  }
}
```



This is intended to be a master script, running other scripts, and possibly serving as a starting point for an eventual paper. 
It also at the moment (3rd of august) serves as a general doc capturing thoughts on the project. 




## Diseases to cover ##
I had a chat with Mark Bronsvoort, and we spent some time brainstorming potentially interesting diseases. IDeally, we would try and create a narrative, where we go from local to global datasets, and using both human and animal diseases. I'm still working on the animal outbreak databases.


### Locally###
 Lyme's disease (George gunn Scotland) 
 Size of TB outbreaks in Scotland - Issue with defining outbreak size here - number of cattles or number of infected herds? 
 
### Regionally (Europe) ###
 E.Coli outbreaks in humans (European)
 
 
### Globally ###
 Cryptosporidium 
 Lumpy skin disease
 African Swine Fever
Dengue - Of high interest as an EID, but has the problem of being endemic... How would we apply EVT to endemic or mixed endemic/epidemic time series? We have a databse of "occurency", but I'm unclear what to make of it. 
Rift Valley fever
Ebola - have identified a database for this - see make-ebola.Rmd
FMD - ew can use the OIE-WAHIS database to get a record of outbreaks. again the distinction between endemic (in Africa) and emerging/epidemic (in e.g. Europe/Asia/Americas), and the question of what to do with that.

### General databases ###
Smith etal 

### References of interest ###
 Woolhouse
 EIDs database 
  "70% of zoonotic diseases being emerging" 
 "Trying to estimate the number of pathogens out there." !!
  


# Summary of current data #

First, run the scripts in the separate documents to create datasets of interest
```{r databases}
library(knitr)
source(file.path(script.path,"make-ebola.R"))
source(file.path(script.path,"make-smith-multipathogen-db.R"))
```


Then, lets do some descriptive plotting

```{r descriptive}
hist(ebola_outbreaks$OB_CASE)
hist(smith_cholera$Total.Cases)
hist(smith_meningitis_bact$Total.Cases)
hist(smith_meningitis_viral$Total.Cases)
hist(smith_nipah$Total.Cases)
hist(smith_plague$Total.Cases)
hist(smith_wnv$Total.Cases)
hist(smith_rvf$Total.Cases)
hist(smith_tularemia$Total.Cases)

```
