---
title: "make-dengue"
author: "Gustaf Rydevik"
date: "3 August 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description

The following data set is downloaded from https://figshare.com/articles/Dengue_Occurrence_12122013.xlsx/1035043

quoting: 

A global geographic database of human dengue virus occurrence. The database comprises occurrence data linked to point or polygon locations, derived from peer-reviewed literature and case reports as well as informal online sources. Entries date from 1960 to 2012.

OCCURRENCE_ID: Unique identifier for each occurrence in the database after temporal and locational standardisation.

SOURCE_TYPE: Manual entry vs. HealthMap

LOCATION_TYPE: Whether the record represents a point or a polygon location

ADMIN_LEVEL: The administrative level which the record represents when the location type is a polygon. Values are 0 (national), 1 (state or province), 2 (district), and -999 when the location type is a point.

GAUL_AD0: The country-level global administrative unit layer (GAUL) code (see http://www.fao.org/geonetwork) which identifies the Admin-0 level occurrences as well as the Admin-0 polygon within which any smaller polygons and points lie.

GAUL_AD1: The first-level GAUL code which identifies the Admin-1 level occurrences as well as the Admin-1 polygon within which any smaller polygons and points lie. Values of -999 are assigned when the polygon was Admin-0 level.

GAUL_AD2: The second-level GAUL code which identifies the Admin-2 level occurrences as well as the Admin-2 polygon within which any points lie. Values of -999 are assigned when the polygon was Admin-0 or Admin-1 level.

POINT_ID: The unique identifier which was assigned to points falling within the same 5km x 5km pixel.

UNIQUE_LOCATION: A unique identifier created for all locations (both points and polygons) based upon the point IDs and the GAUL codes.

X: The longitudinal coordinate of the point or polygon centroid (WGS1984 Datum).

Y: The latitudinal coordinate of the point or polygon centroid (WGS1984 Datum).

YEAR: The year of the occurrence.

COUNTRY: The name of the country within which the occurrence lies.

REGION: The region within which the occurrence lies – values are Asia, Oceania (includes Australia), Africa (includes the Arabian peninsula), and Americas.

```{r read_dengue}
autolib<-function(Package,mirror="http://stat.ethz.ch/CRAN",...){
  if(!suppressWarnings(require(as.character(substitute(Package)),
                               character.only=TRUE,quietly=TRUE,...))){
    Package<-as.character(substitute(Package))
    install.packages(Package,repos=mirror,...)
    library(Package,character.only=TRUE)
  }
}
autolib(readxl)
data.path<-"./outbreak-data/dengue"
dengue_occurence_database<-read_excel(file.path(data.path,"Dengue_Occurrence_12122013.xlsx"))
dengue_occurence_year_country<-dengue_occurence_database%>%group_by(YEAR,COUNTRY)%>%summarise(n.occurence=n())

```
