---
title: "make-smith-multipathogen-db"
author: "Gustaf Rydevik"
date: "3 August 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description

The following data set is downloaded from http://ramachandran-data.brown.edu/datarepo/request.php?request=explorePublicStudy&StudyID=6&instit=BROWN,
and linked to an article published at http://rsif.royalsocietypublishing.org/content/11/101/20140950 . 

quoting: This is a README file describing the outbreak data referenced in the Smith et al. paper. These data are available in the file SmithEtAl_outbreakdata.xlsx. 

Below is a list of the column headers and a brief description:

UID:	A unique numerical ID assigned to each outbreak
Disease:	The disease name, as listed in GIDEON
Country:	The nation where the outbreak occurred, as listed in GIDEON
Year:	The year of the outbreak, or the first year if a range of years was given
Total Cases:	Total number of cases for the given outbreak. (Outbreaks with no case data listed have a value of 9999.)
Transmission Type:	0=Non-vector borne, 1=Vector Borne
Host Type:	0=Zoonotic, 1=Human-specific
Pathogen Taxonomy:	1=bacterium, 2=virus, 3=protozoon, 4=parasite (i.e. helminth), 5=fungus, 6=alga

For more details, please see the manuscript.


```{r read_smith}
script.path<-"./scripts"
source(file.path(script.path,"make-smith-multipathogen-db.R"))
pextRemes(fevd((smith_chikungunya%>%group_by(Year)%>%summarise(max.cases=max(Total.Cases)))$max.cases),1e8)
```
